---
title: "Algal biomass variability calculation and plot illustration"
author: "Thomas Van Zuiden"
date: "May 25, 2017"
output: html_document
---
#### Load required packages
```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(zoo)
library(cowplot)
```

### Import data and see how it looks (determine its structure)
```{r echo=FALSE}
setwd("C:\\Users\\adminuser\\Desktop\\EVERYTHING THOMAS\\gitprojects\\30yr_Algae_Variability_Dorset")
```
```{r results="hide"}
dat1<-read.csv("01RELdorset.csv")
dat2<-read.csv("02ABSdorset.csv")
names(dat1)
names(dat2)
```
```{r results="hide"}
al_rel <- dat1[,c(1,4:10)]
al_abs <- dat2[,c(3:9)]
names(al_rel)
names(al_abs)
```

### 5 year rolling standard deviation window for relative and absolute algal biomasses
```{r}
df <- data.frame() # Create empty data frames
df1 <- data.frame()
for (i in unique(al_rel$Lake)){
  relsd_i <- rollapply(al_rel[al_rel$Lake==i,3:8],5,sd,fill=NA,by.column=TRUE)
  abssd_i <- rollapply(al_abs[al_abs$Lake==i,2:7],5,sd,fill=NA,by.column=TRUE)
  df <- rbind(df,relsd_i)
  df1<- rbind(df1, abssd_i)
}
```

### Creating a new data frame with the algae biomasses and their respective 5 year variability windows
```{r results="hide"}
new_dat <- data.frame(al_rel,df,al_abs,df1)
names(new_dat)
new_dat <- new_dat[, -c(15), drop=FALSE] ### Dropping the extra "year" column
head(new_dat)
```

### 2 by 2 grid of chrysophyte biomasses plots and their 5 year variabilities in 9 Dorset lakes

#### P.S. if anyone knows a way to write cleaner ggplot2 code, I'd love to hear it
```{r warning=FALSE, message=FALSE}
A <- ggplot(data=new_dat, aes(x=Year, y=Chrysophytes_REL))+geom_line(aes(colour=Lake), size=1.1)+
  scale_color_manual(values=c(6,"red2", "orange","yellow2","green","darkgreen","cyan","blue",1,8))+
  xlab("Year")+ylab("Relative Chrysophyte Biomass")+ylim(c(0,80))+
  theme_grey(base_size = 11, base_family = "Helvetica")+theme(legend.position="none")

C <- ggplot(data=new_dat, aes(x=Year, y=Chrysophytes_REL.1))+geom_line(aes(colour=Lake), size=1.1)+
  scale_color_manual(values=c(6,"red2", "orange","yellow2","green","darkgreen","cyan","blue",1,8))+
  xlab("Year")+ylab("5Yr Relative Variability")+ylim(c(0,30))+
  theme_grey(base_size = 11, base_family = "Helvetica")+theme(legend.position="none")

B <- ggplot(data=new_dat, aes(x=Year, y=Chrysophytes_BV))+geom_line(aes(colour=Lake), size=1.1)+
  scale_color_manual(values=c(6,"red2", "orange","yellow2","green","darkgreen","cyan","blue",1,8))+
  xlab("Year")+ylab("Total Chrysophyte Biovolume")+ylim(c(0,3600))+
  theme_grey(base_size = 11, base_family = "Helvetica")+theme(legend.position="none")

D <- ggplot(data=new_dat, aes(x=Year, y=Chrysophytes_BV.1))+geom_line(aes(colour=Lake), size=1.1)+
  scale_color_manual(values=c(6,"red2", "orange","yellow2","green","darkgreen","cyan","blue",1,8))+
  xlab("Year")+ylab("5Yr Biovolume Variability")+ylim(c(0,1500))+
  theme_grey(base_size = 11, base_family = "Helvetica")+theme(legend.position="none")

prow <- plot_grid( A + theme(legend.position="none",axis.text.x=element_blank(),
                             axis.title.x=element_blank()),
                   C + theme(legend.position="none",axis.text.x=element_blank(),
                             axis.title.x=element_blank()),
                   B + theme(legend.position="none"),
                   D + theme(legend.position="none"),
                   align = 'vh',
                   hjust = -1,
                   nrow = 2
)

legend <- get_legend(A + theme(legend.position="left"))
p <- plot_grid( prow, legend, rel_widths = c(3,0.85))
p
```

### create the same plots for the other species of Algae

### See the code from the "variability_codes_plots.R" directory for a 3x3 grid of plots alternative to the 2x2 grid shown above